---
title: "HW2 Prince George's County Crime"
---

In this assignment, we will be working with a dataset reporting the number of crimes by month in Prince George's County, Maryland, for every month in 2024. 

### Task 1: Load the Tidyverse

You will need to load the tidyverse library

```{r}
library(tidyverse)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/pg_crime_totals_2024.csv. You will create a dataframe called `pg_crime` that contains the data from this URL.

### Task 2: Load the Data and Describe It

```{r}
pg_crime <-read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/pg_crime_totals_2024.csv")
```
### Task 3: Describe the Data

You should examine that dataframe carefully by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the categories of incidents. From a reporting standpoint, are any of the categories more important than others? Why or why not? Do any of them make the job of summarizing the data more difficult? Why or why not?

The month of October was the worst for accidents in the county. This could be because car accidents could happen more frequently on dates that surround Halloween and Halloweekend/week due to young people celebrating the holiday under the influence. 

The number of shootings increased and stayed consistent during the summer months. This could be because of the heat making others more reactive and angry. It is interesting though that throughout July and August the number of shootings were consistently 21. 

It is also interesting that only about 10% of the cars that were stolen were recovered each month. 

I think that Homicide, shootings and sexual offences are some of the most important categories to look at as it can highlight 

### Task 4: Calculating Averages

Add up the total number of incidents per month for each category using `mutate()`, calling the new column `total_incidents`. Then calculate the average number of incidents per month for each category of crime using the total number of months, calling the new column `average_incidents`. Assign the result to a new dataframe called `pg_crime_averages`.

```{r}
pg_crime_averages <- pg_crime |> mutate(((total_incidents = Jan + Feb + Mar + Apr + May + Jun + Jul + Aug + Sep + Oct + Nov + Dec)/12))
```

### Task 5: Describe the Averages

Examine the new dataframe, `pg_crime_averages`, by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the average number of incidents per month for each category. Which categories have the highest average number of incidents per month? Which have the lowest? What does this tell you about the pattern of crime in Prince George's County? Are there categories that you have questions about? Why or why not?

The category with the highest average number was Accidents which was 481 while the lowest was Robbery Residential which is n/a. This just shows us that the number of accidents happens more than residential robberies in the county. I wonder how accurate the sex offense numbers are as lots of these cases go unreported. 


### Task 6: Testing a Theory

One theory about crime is that it is more likely to occur in the summer months. Test this theory by creating three new columns using `mutate()` that calculate the percentage change compared to the average for each category of crime for the months of June, July, and August. Call the new columns `june_change`, `july_change`, and `august_change`. Assign the result to a new dataframe called `summer_crime_averages`.

```{r}
summer_crime_averages <-pg_crime |>  
  mutate( june_change = ((127.5555555556-123.65277778333) / 123.65277778333) * 100, july_change = ((126.7222222222-123.65277778333) / 123.65277778333) * 100, august_change = ((124.7222222222222-123.65277778333) / 123.65277778333) * 100)
```


### Task 7: Describe the Summer Crime Averages

Look at the `summer_crime_averages` dataframe by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the percentage change in the average number of incidents per month for each category of crime for the months of June, July, and August. Do you see any patterns? What does this tell you about the theory that crime is more likely to occur in the summer months?

The august percent change is lower than the other two months. The change gets lower as the months go by. It shows that in June it is more likely to occur than in the other two months. 

### Task 8: What's the Story?

Write up a lede and nutgraf of a story focused on the crime data in Prince George's County. It could be from any of the dataframes you've created in this assignment. The lede should be a single sentence that summarizes the most important information in the data. The nutgraf should be a paragraph that provides more context and explains why the data is important.

Residential Robbery occurs consistently in the summer months than in the other months of the year. 

From June to August the number of residential robberies was 3 each month. This number is stable which concludes that there was not an influx or dip in the number of residential robberies during those months. This data is important as it represents that the county has done a good job at keeping the number of residential robberies low and consistent. Later, during the following months, the number of robberies slowly decreases. 

### Task 9: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.
