---
title: "HW8_MandMs"
---

In this assignment, you'll be picking up where you left off with the M&Ms activity. You'll load the data and perform some basic statistical analysis. Follow the steps below to complete the assignment.

### Task 1: Load the tidyverse

```{r}
library(tidyverse)
```

### Task 2: Load the M&Ms dataset from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/mandms.csv
```{r}
mm_data<-read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/mandms.csv")
```


### Task 3: Sample 5 rows of data from the dataset
```{r}
sample_5 <- mm_data |>sample_n(5)

```


### Task 4: Calculate the mean age and standard deviation of the sample, using either the red or black number of M&Ms

```{r}
sample_5 |> summarize(mean_red = mean(red), sd_red = sd(red))
```



### Task 5: Calculate the mean age and standard deviation of the population, using the same measure as in Task 4

```{r}
mm_data |> summarize(mean_red = mean(red), sd_red = sd(red))
```



### Task 6: Compare the sample and population statistics

The mean of the sample is 70.8 while the mean of the entire population is 72.98 they are off by about 2.18 percentage points. While the standard deviation of the sample is 4.4 and the standard deviation of the population is 16.4 which is a difference of 12 percentage points. 
### Task 7: Describe the results

Write a few sentences about what you found in the data. What does this tell you about the M&Ms dataset and about the act of random sampling?

The differences between the population and the sample mean and standard deviation shows us that a sample size of five does not accurately represent the entire population of 100. The data from the random sample is fairly centered as the mean's are close to each other but the standard deviations are so drastically different which shows how spread out the data is from each other in the total than the sample. 

When finished, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
